---
title: "Model"
format: html
editor_options: 
  chunk_output_type: inline
---

# Predictive Model

### Libraries

```{r}
library(tidyverse)
library(plotrix)
library(corrplot)
library(pROC)
library(doParallel)
library(rpart)
library(smotefamily)
library(xgboost)
library(tidymodels)
library(finetune)
library(caret)
library(themis)
library(fastDummies)
library(naniar)

```

## Loading Data

```{r}

dat <- read_csv("returns_train.csv")
```

### Viz Miss - Visualizing missing data

```{r}
dat %>%
  select(where(~any(is.na(.)))) %>%
  vis_miss() +
  labs(title = "Missing Data") +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, margin = margin(t = 50), size = 11),
    plot.title = element_text(margin = margin(b=10),
                              size = 25,
                              hjust = 0.5))
```

### Balance column chart

```{r}

props <- prop.table(table(dat$returned))

barplot(props, 
        ylim = c(0, 1),
        col = c("hotpink", "skyblue"),
        main = "Returned Class Balance",
        ylab = "Proportion")
text(x = c(0.7, 1.9),
     y = props + 0.05, 
     labels = paste0(round(props *100, 1), "%"),
     cex = 1.5)  

```

### Violin plot - shows outliers and distribution

```{r}

```

### Scatterplot Matrix

```{r}

```

## Cleaning Data

```{r}

data_cleaning <- function(df) {
  
  # Remove unique identifier 
  df$transaction_id <- NULL
  
  # Make transaction_timestmap a date format
  if("transaction_timestamp" %in% names(df)) {
    df$transaction_timestamp <- lubridate::ymd_hms(df$transaction_timestamp)
  }
  
  # Make customer_age absolute factors
  if("customer_age" %in% names(df)) {
    df$customer_age <- abs(df$customer_age)
  }
  
  # Fix payment method 
  if("payment_method" %in% names(df)) {
    df$payment_method <- tolower(trimws(df$payment_method))
    df$payment_method <- recode(df$payment_method,"credit card - visa" = "visa")
  }
  
  return(df)
}

dat <- data_cleaning(dat) 
```

### Correlation Plot

```{r}

cor_matrix <- cor((dat %>%
             select(price, discount_pct, customer_age, account_age_days, returned)), use = "pairwise.complete.obs")

corrplot(cor_matrix,
         method = "color",
         type = "upper",
         addCoef.col = "black",
         tl.col = "black",
         tl.srt = 45,
         diag = FALSE,
         title = "\n\nCorrelation Matrix of Numeric Features",
         addgrid.col = "gray80",
         mar = c(0, 0, 1, 0))


```

## Feature Variables

```{r}

feature_variables <- function(df) {
  
  if("applied_promo_codes" %in% names(df)) {
    df$promo_winter50 <- as.numeric(str_detect(df$applied_promo_codes, "WINTER50"))
    df$promo_freeship <- as.numeric(str_detect(df$applied_promo_codes, "FREESHIP"))
    df$promo_save20 <- as.numeric(str_detect(df$applied_promo_codes, "SAVE20"))
    df$promo_newuser <- as.numeric(str_detect(df$applied_promo_codes, "NEWUSER"))
    
    df$promo_count <- ifelse(is.na(df$applied_promo_codes), 0, 
                             str_count(df$applied_promo_codes, ",") + 1)
    
    df <- df %>%
      mutate(across(starts_with("promo_"), ~replace_na(., 0)))
    
    df$applied_promo_codes <- NULL
  }
  
  
  # Extract time, day, month, year of timestamps 
  df$time <- as.integer(format(df$transaction_timestamp, "%H%M%S"))
  df$day <- as.integer(format(df$transaction_timestamp, "%d"))
  df$month <- as.integer(format(df$transaction_timestamp, "%m"))
  df$year <- as.integer(format(df$transaction_timestamp, "%Y"))
  df$transaction_timestamp <- NULL
  
  return(df)
  
}

dat <- feature_variables(dat)
```

## Factor Variables

```{r}
make_factors <- function(df) {
  
  # Make variables into factors 
  df$marketing_channel <- as.factor(df$marketing_channel)
  df$product_category <- as.factor(df$product_category)
  df$product_subcategory <- as.factor(df$product_subcategory)
  df$payment_method <- as.factor(df$payment_method)
  df$guest_checkout <- as.factor(df$guest_checkout)
  df$loyalty_tier <- factor(df$loyalty_tier, 
                          levels = c("None", "Bronze", "Silver", "Gold", "Platinum"))
  if("returned" %in% names(df)) {
     df$returned <- factor(df$returned)
  }
  
  df$zip_code <- as.factor(df$zip_code)
  
  return(df)
}

dat <- make_factors(dat)
```

## Removing low importance variables

```{r}

remove_var <- function(df) {
  df$guest_checkout <- NULL
  df$loyalty_tier <- NULL
  df$marketing_channel <- NULL
  df$payment_method <- NULL
  df$promo_winter50 <- NULL
  df$promo_save20 <- NULL
  df$promo_newuser <- NULL
  df$month <- NULL 
  #df$promo_count <- NULL
  #df$customer_age <- NULL
  #df$day <- NULL

  return(df)
}

dat <- remove_var(dat)
```

## Split the data

```{r}
set.seed(42)

train <- createDataPartition(dat$returned, p = 0.90, list = FALSE)
datTrain <- dat[train, ]
datTest <- dat[-train, ]

dim(datTrain)
dim(datTest)
```

## Random Under Sampling

```{r}

prop.table(table(datTrain$returned))

recipe_obj <- recipe(returned~., data = datTrain) %>%
  step_downsample(returned, under_ratio = 2) %>%
  prep()

datTrain <- recipe_obj %>%
  bake(new_data = NULL)

print(prop.table(table(datTrain$returned)))
```

## XGBoost Modeling

### Define matrices

```{r}

# Define matrices 
train_x <- data.matrix(datTrain[, -which(names(datTrain) == "returned")])
train_y <- as.numeric(datTrain$returned) - 1

test_x <- data.matrix(datTest[, -which(names(datTest) == "returned")])
test_y <- as.numeric(datTest$returned) - 1

xgb_train <- xgb.DMatrix(data = train_x, label = train_y)
xgb_test <- xgb.DMatrix(data= test_x, label = test_y)
```

### Fit XGBoost

```{r}
# Define watchlist
watchlist = list(train = xgb_train, test = xgb_test)

# Fit model
xgb_model = xgb.train(
  params = list(
    learning_rate = 0.03, 
    max_depth = 4, 
    objective = "binary:logistic",
    eval_metric = "auc",
    subsample = 0.5,
    min_child_weight = 21,
    colsample_bytree = 0.2,
    gamma = 3
  ),
  data = xgb_train, 
  evals = watchlist, 
  nrounds= 3000,
  early_stopping_rounds = 300,
  print_every_n = 100
)

eval_log <- attributes(xgb_model)$evaluation_log
best_round <- which.max(eval_log$test_auc)
print(paste("Best round:", best_round,
            "Test AUC:", eval_log$test_auc[best_round]))

```

### Predict

```{r}
pred_xgb <-predict(xgb_model, newdata = xgb_test, iterationrange = c(1, best_round +1))

roc_xgb <- roc(test_y, pred_xgb, quiet = TRUE)
print(auc(roc_xgb))
```

## Testing Data

### Testing data set up

```{r}
kaggle <- read_csv("returns_test.csv")
test_id <- kaggle$transaction_id

# Cleaning and processing kaggle data

kaggle <- data_cleaning(kaggle)
kaggle <- feature_variables(kaggle)
kaggle <- make_factors(kaggle)
kaggle <- remove_var(kaggle)

# Define matrices

names(kaggle)
kaggle_x <- data.matrix(kaggle)

xgb_kaggle_test <- xgb.DMatrix(data = kaggle_x)

```

### Predict on Test Data

```{r}
kaggle_preds <- predict(xgb_model, 
                        newdata = xgb_kaggle_test,
                        iterationrange = c(1, best_round + 1))
```

### Make Submission File

```{r}
submission <- data.frame(transaction_id = test_id, 
                         returned = kaggle_preds)

write.csv(submission, "submission_41.csv", row.names = FALSE)
```

## Importance Variable

```{r}
importance_matrix <- xgb.importance(model = xgb_model)

xgb.ggplot.importance(importance_matrix) +
  ggtitle("XGBoost Variable Importance") +
  theme_minimal() 
```

## CV XGBoost

### Fit CV xgboost

```{r}
# 1. Run the CV to find the 'True' AUC and the best number of rounds
cv_model <- xgb.cv(
  params = list(
    learning_rate = 0.03, 
    max_depth = 4,
    objective = "binary:logistic",
    eval_metric = "auc",
    subsample = 0.5,
    colsample_bytree = 0.2,
    min_child_weight = 21,
    gamma = 3
  ),
  data = xgb_train,        # Use your 66/33 undersampled DMatrix here
  nfold = 5,               # 5 folds is standard for Kaggle
  nrounds = 3000,          
  early_stopping_rounds = 300,
  print_every_n = 100,
  stratified = TRUE,       # CRITICAL: Keeps the 66/33 balance in every fold
  verbose = 1
)

# 2. Extract the log (it's a data.table)
cv_log <- cv_model$evaluation_log

# 3. Find the row index where the average test AUC was highest
best_round <- which.max(cv_log$test_auc_mean)

# 4. Double check it
print(paste("The best round is:", best_round))

# 3. Look at your 'True' Expected Score
# This 'test_auc_mean' is usually much closer to the Kaggle Leaderboard
final_cv_auc <- max(cv_model$evaluation_log$test_auc_mean)
print(paste("Expected Kaggle AUC:", round(final_cv_auc, 4)))

final_xgb_model <- xgb.train(
  params = list(
    learning_rate = 0.03, 
    max_depth = 4,
    objective = "binary:logistic",
    eval_metric = "auc",
    subsample = 0.5,
    colsample_bytree = 0.2,
    min_child_weight = 21,
    gamma = 3
  ),
  data = xgb_train, # Your full DMatrix
  nrounds = best_round
)

```

### Predict CV XGboost

```{r}
# Predict
pred_xgb_cv <-predict(final_xgb_model, newdata = xgb_test)

roc_xgb_cv <- roc(test_y, pred_xgb_cv, quiet = TRUE)
print(auc(roc_xgb_cv))
```

## XGBoost Autotuning

### Prepare data for tidymodels

```{r}

# Recode returned to "Yes"/"No" for tidymodels
# train_balanced$returned is already a factor (0/1 levels from make_factors),
# so we check which level corresponds to 1

xgb_data <- train_balanced %>%
  mutate(returned = factor(ifelse(returned == 1, "Yes", "No"),
                           levels = c("No", "Yes")))

xgb_data_test <- datTest %>%
  mutate(returned = factor(ifelse(returned == 1, "Yes", "No"),
                           levels = c("No", "Yes")))
```

### Model Specification

```{r}
xgb_spec <- boost_tree(
  trees          = tune(),
  tree_depth     = tune(),
  learn_rate     = tune(),
  loss_reduction = tune(),
  min_n          = tune(),
  sample_size    = tune(),
  mtry           = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")
```

### CV Folds and Search Grid

```{r}

set.seed(42)
cv_folds <- vfold_cv(xgb_data, v = 5, repeats = 3, strata = returned)

set.seed(42)
xgb_grid <- grid_random(
  trees(range          = c(100, 1000)),
  tree_depth(range     = c(3, 10)),
  learn_rate(range     = c(-3, -1)),    # log10 scale: 0.001 to 0.1
  loss_reduction(range = c(0, 10)),
  min_n(range          = c(1, 20)),
  sample_prop(range    = c(0.5, 1.0)),
  mtry(range           = c(5, ncol(xgb_data) - 1)),
  size = 100
)
```

### Workflow

```{r}
xgb_workflow <- workflow() %>%
  add_model(xgb_spec) %>%
  add_formula(returned ~ .)
```

### Run Tuning

```{r}
#| cache: true
#| label: tuning

cl <- makePSOCKcluster(detectCores() - 1)
registerDoParallel(cl)

start <- Sys.time()
set.seed(42)
xgb_tuned <- tune_grid(
  xgb_workflow,
  resamples = cv_folds,
  grid      = xgb_grid,
  metrics   = metric_set(roc_auc),
  control   = control_grid(verbose = TRUE, allow_par = TRUE)
)
stopCluster(cl)

print(Sys.time() - start)
```

### Tuning Results

```{r}
# Top 5 parameter combinations
show_best(xgb_tuned, metric = "roc_auc", n = 5)
```

```{r}
# Visualise tuning performance across hyperparameters
autoplot(xgb_tuned)
```

### Fit Final Model with Best Paramemete

```{r}
best_params <- select_best(xgb_tuned, metric = "roc_auc")
print(best_params)

xgb_final_workflow <- xgb_workflow %>%
  finalize_workflow(best_params)

xgb_final_fit <- fit(xgb_final_workflow, data = xgb_data)
```

### Evaluate on test set

```{r}
xgb_final_probs <- predict(xgb_final_fit,
                           new_data = xgb_data_test,
                           type = "prob")

roc_result <- roc(
  response  = as.numeric(xgb_data_test$returned == "Yes"),
  predictor = xgb_final_probs$.pred_Yes,
  quiet     = TRUE
)
print(auc(roc_result))
```

## Importance Variable

```{r}
importance_matrix <- xgb.importance(model = xgb_model)

xgb.ggplot.importance(importance_matrix) +
  ggtitle("XGBoost Variable Importance") +
  theme_minimal() 
```
